### 2019-02-20 Paper1/colloc_ipm.tex
added file structure

\section{Collocation accelerated IPM for steady state problems}