### 2019-02-21 Paper1/summary_outlook.tex
continued

\section{Summary and outlook}
\label{sec:summary_outlook}