### 2019-11-22 Paper1/main.tex
new abstract

\documentclass[3p]{elsarticle}

% Packages and macros go here
\usepackage{bm}
\usepackage[parfill]{parskip}
\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{color}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{array}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

\ifpdf
  \DeclareGraphicsExtensions{.eps,.pdf,.png,.jpg}
\else
  \DeclareGraphicsExtensions{.eps}
\fi

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}[theorem]{Remark}

\DeclareMathOperator*{\argmin}{arg\,min}

\makeatletter
\newsavebox\myboxA
\newsavebox\myboxB
\newlength\mylenA

\newcommand*\xoverline[2][0.75]{%
    \sbox{\myboxA}{$\m@th#2$}%
    \setbox\myboxB\null% Phantom box
    \ht\myboxB=\ht\myboxA%
    \dp\myboxB=\dp\myboxA%
    \wd\myboxB=#1\wd\myboxA% Scale phantom
    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
    \addtolength\mylenA{-\the\wd\myboxB}%
    \ifdim\wd\myboxB<\wd\myboxA%
       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
    \else
        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
    \fi}
\makeatother

\journal{arXiv.org}

%strongly recommended
%\numberwithin{theorem}{section}

% Declare title and authors, without \thanks
\newcommand{\TheTitle}{Intrusive acceleration strategies for Uncertainty Quantification for hyperbolic systems of conservation laws}

\date{\today}

% Sets running headers as well as PDF title and authors
%\headers{\RunningTitle}{\TheAuthors}

\usepackage{amsopn}
\DeclareMathOperator{\diag}{diag}
%\DeclareMathOperator{\argmin}{arg\min}

% FundRef data to be entered by SIAM
%<funding-group>
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\def\lambdabar{\xoverline{\bm{\lambda}}}
\def\lambdabarjn{\lambdabar_j^n}

\def\remja#1{{\color{red}{\bfseries \fbox{Remark {\color{red}Jannick}}}\enspace #1}}
\def\frama#1{{\color{red}{\bfseries \fbox{{\color{blue}Martin}}}\enspace #1}}
\newcommand{\comment}[1]{{\textcolor{blue}{#1}}}

\begin{document}

\begin{frontmatter}

\title{\TheTitle}

\author[adressJonas]{Jonas Kusch}
\author[adressJannick]{Jannick Wolters}
\author[adressMartin]{Martin Frank}

\address[adressJonas]{Karlsruhe Institute of Technology, Karlsruhe,
    jonas.kusch@kit.edu}
\address[adressJannick]{Karlsruhe Institute of Technology, Karlsruhe, jannick.wolters@kit.edu}
\address[adressMartin]{Karlsruhe Institute of Technology, Karlsruhe, martin.frank@kit.edu}

\begin{abstract}
Methods for quantifying the effects of uncertainties in hyperbolic problems can be divided into intrusive and non-intrusive techniques. Non-intrusive methods allow the usage of a given deterministic solver in a black-box manner, while being embarrassingly parallel. However, abandoning the idea of intrusively modifying a given solver takes away the ability to use several numerical acceleration tools, which have an intrusive nature. One example is the use of adaptivity for which intrusive methods provide a convenient framework. Moreover, intrusive methods are expected to reach a given accuracy with a smaller number of unknowns than non-intrusive techniques. This effect is amplified in settings with high dimensional uncertainty where the number of polynomials is smaller than the number of points in a sparse grid to achieve a certain total degree. A downside of intrusive methods is however the need to guarantee hyperbolicity of the moment system as well as numerical costs: Compared to stochastic-Galerkin (SG), the Intrusive Polynomial Moment (IPM) method maintains hyperbolicity, but requires solving an optimization problem in every spatial cell and every time step. \\
In this work, we propose acceleration techniques for intrusive methods and study their advantages and shortcomings compared to the non-intrusive stochastic-Collocation method. When solving steady problems with IPM, the numerical costs arising from repeatedly solving the IPM optimization problem can be reduced by borrowing ideas from PDE-constrained optimization. Integrating the iteration from the numerical treatment of the optimization problem into the moment update reduces numerical costs, while preserving local convergence. Additionally, we propose an adaptive implementation and an efficient parallelization of the IPM method. The effectiveness of the proposed strategies is demonstrated for multi-dimensional uncertainties in fluid dynamics applications. In our numerical experiments, we observe the expected behavior of requiring a smaller number of unknowns to achieve a given accuracy when using intrusive methods. Using the proposed acceleration techniques, our implementation reaches a given accuracy faster than stochastic-Collocation.
\end{abstract}

\begin{keyword}
uncertainty quantification, conservation laws, hyperbolic, intrusive, stochastic-Galerkin, Collocation, Intrusive Polynomial Moment Method
\end{keyword}

\end{frontmatter}

\input{introduction}
\input{framework}
%\section{Strategies for steady problems}
%\input{colloc_ipm}
\input{one_shot_ipm}
\input{adaptivity}
\input{implementation}
\input{results}
\input{summary_outlook}

\newpage
\bibliographystyle{unsrt}  
\bibliography{references}
\input{appendix}
\end{document}