### 2020-05-25 Paper1/ResponseToReviewers/CRediT.tex
Revised Submission JCP

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand*\xoverline[2][0.75]{%
    \sbox{\myboxA}{$\m@th#2$}%
    \setbox\myboxB\null% Phantom box
    \ht\myboxB=\ht\myboxA%
    \dp\myboxB=\dp\myboxA%
    \wd\myboxB=#1\wd\myboxA% Scale phantom
    \sbox\myboxB{$\m@th\overline{\copy\myboxB}$}%  Overlined phantom
    \setlength\mylenA{\the\wd\myboxA}%   calc width diff
    \addtolength\mylenA{-\the\wd\myboxB}%
    \ifdim\wd\myboxB<\wd\myboxA%
       \rlap{\hskip 0.5\mylenA\usebox\myboxB}{\usebox\myboxA}%
    \else
        \hskip -0.5\mylenA\rlap{\usebox\myboxA}{\hskip 0.5\mylenA\usebox\myboxB}%
    \fi}
\makeatother
\usepackage{bm}

\title{Reply to Reviewers}
\author{Jonas Kusch, Jannick Wolters and Martin Frank}
\date{\today}                                           % Activate to display a given date or no date

\setlength{\parindent}{0pt}

\begin{document}
\textbf{Jonas Kusch:} Conceptualization, Software, Validation, Formal analysis, Investigation, Writing - Original Draft, Writing - Review \& Editing \\
\textbf{Jannick Wolters:} Software, Validation, Investigation,  Data Curation, Writing - Original Draft, Visualization \\
\textbf{Martin Frank:} Methodology, Writing - Original Draft, Supervision, Funding acquisition  
\end{document}