### 2020-07-03 Paper1/SignificanceAndNovelty/highlights.tex
Paper version elsevier

 %% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.
%Version for spanish users, by dgarhdez

\documentclass[10pt,a4paper]{article}      % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')
%\usepackage[english]{babel}

%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}

\subsection*{Highlights}
- intrusive methods for uncertainty quantification have high numerical costs\\
- adaptivity and the One-Shot approach can be used to accelerate intrusive methods\\
- proposed techniques are studied for uncertainties up to order three\\
- acceleration techniques allow intrusive methods to compete with collocation methods
\end{document}

%% end of file `template.tex'.